### 2021-03-19 paper3/appendix.tex
Address the remaining feedback

%% Recommended, but optional, packages for figures and better typesetting:
%\usepackage{microtype}
%\usepackage{graphicx}
%\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{booktabs} % for professional tables
%\usepackage{amsfonts}       % blackboard math symbols
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{amsthm}
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{xcolor}         %for coloring the text on colorMNIST
%\usepackage{comment}
%\usepackage{enumitem}
%\newtheorem{prop}{Proposition}
%\newcommand{\Xcal}{\mathcal{X}}
%\newcommand{\Bcal}{\mathcal{B}}
%\newcommand{\Ycal}{\mathcal{Y}}
%\newcommand{\Scal}{\mathcal{S}}
%\newcommand{\Lcal}{\mathcal{L}}
%\newcommand{\Dcal}{\mathcal{D}}
%\newcommand{\ie}{i.\,e.}
%\newcommand{\Ie}{I.\,e.}
%\newcommand{\eg}{e.\,g.}
%\newcommand{\Eg}{E.\,g.}
%% \newcommand{\xmark}{\text{\ding{55}}}
%\usepackage{pifont}% http://ctan.org/pkg/pifont
%\newcommand{\cmark}{\ding{51}}%
%\newcommand{\xmark}{\ding{55}}%

@@

\section{Appendix}\label{sec:zsf-appendix}
\subsection{Results for Adult Income}\label{sec:adult-results}
\begin{figure*}[p]
    \centering
    \includegraphics[width=\columnwidth]{paper3/figures/adult_partial_acc.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_partial_prr.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_partial_tprr.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_partial_tnrr.pdf}
    \caption{%
    Results for the Adult Income dataset with \emph{subgroup bias},
    for the binary classification task of predicting whether an individual earns $>$\$50,000 with a binary subgrouping based on \emph{gender}.
    \texttt{ERM (LD)} refers to a model based on ERM (empirical risk minimization),
    trained on a \emph{l}abelled \emph{d}eployment set; thus not suffering from bias in the training set.
    \textsc{Top left}: Accuracy.
    \textsc{Top right}: Positive rate ratio.
    \textsc{Bottom left}: True positive rate ratio.
    \textsc{Bottom right}: True negative rate ratio.
    For the \texttt{Ranking} clustering, the clustering accuracy was 69.7\% $\pm$ 0.3\%;
    for \texttt{K-means} it was 43\% $\pm$ 3\%.
    }%
    \label{fig:adult-subgroup-bias}
\end{figure*}
\begin{figure*}[p]
    \centering
    \includegraphics[width=\columnwidth]{paper3/figures/adult_miss_s_acc.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_miss_s_prr.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_miss_s_tprr.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_miss_s_tnrr.pdf}
    \caption{%
    Results for the Adult Income dataset with a \emph{missing subgroup},
    for the binary classification task of predicting whether an individual earns $>$\$50,000 with a binary subgrouping based on \emph{gender}.
    \texttt{ERM (LD)} refers to a model based on ERM (empirical risk minimization),
    trained on a \emph{l}abelled \emph{d}eployment set; thus not suffering from bias in the training set.
    \textsc{Top left}: Accuracy.
    \textsc{Top right}: Positive rate ratio.
    \textsc{Bottom left}: True positive rate ratio.
    \textsc{Bottom right}: True negative rate ratio.
    For the \texttt{Ranking} clustering, the clustering accuracy was 60.4\% $\pm$ 0.8\%;
    for \texttt{K-means} it was 44\% $\pm$ 3\%.
    }%
    \label{fig:adult-missing-subgroup}
\end{figure*}
Figures~\ref{fig:adult-subgroup-bias} and \ref{fig:adult-missing-subgroup} show results from our method on the Adult Income dataset \cite{Dua:2017}.
This dataset is a common dataset for evaluating fair machine learning models. 
Each instance in the dataset is described by $14$ characteristics including gender, education, marital status, number of work hours per week among others, along with a label denoting income level ($\geq$\$50K or not). 
We transform the representation into $62$ real and binary features along with the subgroup label $s$. %%%%%%%%CHECK THESE VALUES%%%%%%%%%%%% 
The dataset is naturally imbalanced with respect to gender: 30\% of the males are labelled as earning more than \$50K per year (high income), while only 11\% of females are labelled as such.
For further details on the dataset construction, see section~\ref{ssec:dataset-construction-adult}.
%
Following standard practice in algorithmic fairness, e.g. \citet{zemel2013learning}, we consider gender to be the subgroup label $s$.

@@

As before, \texttt{Ranking}, \texttt{k-means}, \texttt{No bal.}\ and \texttt{Perfect} refer to our method with different procedures for constructing (approximately) perfect bags.
As baseline methods, we have \texttt{ERM} (standard empirical risk minimization with balanced batches), \texttt{DRO} \citep{HasSriNamLia18}, \texttt{gDRO} \citep{sagawa2019distributionally}
and \texttt{ERM (LD)} which is the same model as \texttt{ERM}, but trained on the labelled deployment set, in addition to the training set.

@@

\paragraph{Adult Income.}\label{ssec:dataset-construction-adult}
For the Adult Income dataset, we do not need to apply any synthetic biasing as the dataset naturally contains some bias wrt $s$. Thus, we instantiate the deployment set as just a random subset of the original dataset. Explicit balancing of the test set is needed to yield meaningful evaluation, however, namely in the penalizing of biased classifiers, but need be taken in doing so. Balancing the test set such that
\begin{align}
    |\{x \in X |s=0, y=0\}| &= |\{x \in X |s=1, y=0\}|    \nonumber\\
    \text{and}~|\{x \in X |s=0, y=1\}| &= |\{x \in X |s=1, y=1\}|
\end{align}
where for both target classes, $y=0$ and $y=1$, the proportions of the groups $s=0$ and $s=1$ are made to be the same, is intuitive, yet at the same time precludes sensible comparison of the accuracy/fairness trade-off of the different classifiers.
Indeed, with the above conditions, a majority classifier (predicting all 1s or 0s) achieves comparable accuracy to the fairness-unaware baselines, while also yielding perfect fairness, by construction.
This observation motivated us to devise an alternative scheme, where we balance the test set according to the following constraints
\begin{align}
    & |\{x \in X |s=0, y=0\}| 
    = |\{x \in X |s=0, y=1\}|  \nonumber \\
    = &|\{x \in X |s=1, y=1\}|
    = |\{x \in X |s=1, y=0\}|~.
 \end{align}
That is, all subsets of $\gS \times \gY$ are made to be equally sized. Under this new scheme the accuracy of the the majority classifier is 50\% for the binary-classification task.

\subsection{Optimization}
\begin{table*}[p]
 \centering
 \caption{Selected hyperparameters for experiments with Coloured MNIST, Adult and CelebA datasets.}
 \label{tab:hparams}
\scalebox{0.65}{
 \begin{tabular}{llll}
 \toprule
 & \textsc{Coloured MNIST} & \textsc{Adult} & \textsc{CelebA}       
 \\ & 2-dig SB / 2-dig MS / 3-dig SB
 \\ \midrule
 Input size  &   $3 \times 32 \times 32$ & $61$ & $3 \times 64 \times 64$ \\  \midrule
 \multicolumn{4}{c}{Autoencoder}                     \\ \midrule
 Levels                      & $4$         & $1$    & $5$\\
 Level depth                 & $2$         & $1$    & $2$\\
 Hidden units / level        & $[32, 64, 128, 256]$ & $[61]$ & $[32, 64, 128, 256, 512]$\\
 Activation                  & GELU        & GELU   & GELU   \\
 Downsampling op.  & Strided Convs. & -- & Strided Convs.\\
 Reconstruction loss         & MSE         & Mixed$^1$  & MSE \\
 Learning rate               & $1 \times 10^{-3}$   & $1 \times 10^{-3}$  & $1 \times 10^{-3}$ \\ \midrule
 \multicolumn{4}{c}{Clustering}                      \\ \midrule
 Batch size                  & $256$      & $1000$  & --\\
 AE pre-training epochs      & $150$        & $100$ & --   \\
 Clustering epochs           & $100$       & $300$  & -- \\
 Self-supervised loss & Cosine + BCE & Cosine + BCE & --\\
 U (for ranking statistics)             & $5$         & $3$     & --    \\   \midrule
 \multicolumn{4}{c}{Distribution Matching}                   \\ \midrule
 Batch size & $1$/$32$/$14$  & $64$   & $32$ \\
 Bag size   & $256$/$8$/$18$ & $32$ & $8$ \\
 Training iterations    & $8\text{k}/8\text{k}/20\text{k}$ & $5\text{k}$ & $15\text{k}$ \\
 Encoding ($z$) size$^2$  & $128$   & $35$  & $128$ \\
 Binarised $z_s$ & {\boxedsymbols ✗}\, / \cmark\, / \cmark & \xmark & \xmark \\
 $y$-predictor weight ($\lambda_1$) & $1$ & $0$ & $1$  \\ 
 $s$-predictor weight ($\lambda_2$) & $1$ & $0$ & $0$  \\ 
 Adversarial weight ($\lambda_3$)   & $1 \times
 10^{-3}$   & $1$   & $1$\\ 
 Stopgradient $\left(\nabla_\theta h_\psi(f_\theta(X^\mathit{dep}))=0\right)$ & \xmark & \cmark & \xmark \\
 \midrule
 \multicolumn{4}{c}{Predictors}   \\ \midrule
 Learning rate  & $3 \times 10^{-4}$ &   $1 \times 10^{-3}$  $ 1 \times 10^{-3}$\\
 \midrule
 \multicolumn{4}{c}{Discriminator}                   \\ \midrule
 Attention mechanism$^3$    & Gated   & Gated & Gated \\
 Hidden units pre-aggregation  & $[256, 256]$  & $[32]$ & $[256, 256]$\\
 Hidden units post-aggregation & $[256, 256]$ & --  & $[256, 256]$ \\
 Embedding dim (for attention) & $32$ & $128$ & $128$ \\
 Activation & GELU & GELU & GELU \\
 Learning rate  & $3 \times 10^{-4}$    & $1 \times 10^{-3}$ & $1 \times 10^{-3}$\\
 Updates / AE update    & $1$  & $3$    & $1$    \\
 \bottomrule
 \multicolumn{4}{l}{\footnotesize $^1$ Cross-entropy is used for categorical features, MSE for continuous features.} \\
  \multicolumn{4}{p{\textwidth}}{\footnotesize $^2$ $|z|$ denotes the combined size of $z_s$ and $z_y$, with the former occupying $\ceil{\text{log}_2(\mathcal{S})}$ dimensions, the latter the remaining } \\
  \multicolumn{4}{p{\textwidth}}{\footnotesize dimensions.} \\
  \multicolumn{4}{p{\textwidth}}{\footnotesize $^3$ The attention mechanism used for computing the sample-weights within a bag. \emph{Gated} refers to gated attention  proposed by
%   \cite{ilse2018attention}, while \emph{SDP} refers to the scaled dot-product attention proposed by \cite{vaswani2017attention}.
  } \\ 
  \multicolumn{4}{p{\textwidth}}{\footnotesize  \cite{ilse2018attention}.} \\ 
 \end{tabular}
}
\end{table*}

@@

Since, by design, we do not have labels for all subgroups the model will be tested on, and bias against these missing subgroups is what we aim to avoid, properly validating, and thus conducting hyperparameter selection for models generally, is not straightforward.
We can use estimates of the mutual information between the learned-representation and $s$ and $y$ (which we wish to minimise w.r.t.\ to the former, maximise w.r.t.\ the latter) to guide the process, though optimizing the model w.r.t.\ to these metrics obtained from only the training set does not guarantee generalization to the missing subgroups.
We can, however, additionally measure the entropy of the predictions on the encoded test set and seek to maximise it across all samples, or alternatively train a discriminator of the same kind used for distribution matching as a measure the shift in the latent space between datasets.
We use the latter approach (considering, the learned distance between subspace distributions, accuracy, and reconstruction loss) to inform an extensive grid-search over the hyperparameter space of our model.

@@

In addition to the losses stated in the distribution matching objective, $\mathcal{L}$, in the main text, we also regularise the encoder by the $\ell^2$ norm of its embedding, multiplied by a small pre-factor, finding this to work better than more complex regularization methods, such as spectral normalization \citep{miyato2018spectral}, for stabilizing training.

\subsection{Visualizations of results}\label{sec:qual-results}
Figures~\ref{fig:attn_maps} and \ref{fig:cmnist-recon} show some additional visualizations for our results.
For details, see the captions.
\begin{figure*}[tb]
  \centering
    % \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=0.9\textwidth]{paper3/figures/celeba_attn_map.png}
    % \end{subfigure}
    % \hfill
    % \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=0.9\textwidth]{paper3/figures/cmnist_attn_map.png}
    % \end{subfigure}
  \caption{
    Example sample-wise attention maps for bags of CelebA (left) and Coloured MNIST (right) images sampled from a balanced deployment set. The training set is biased according to the SB setting where for CelebA ``smiling females'' constitute the missing source and for Coloured MNIST {\color{purple}purple} fours constitute the missing source. The attention weights are used during the discriminator's aggregation step to compute a weighted sum over the bag. The attention-weight assigned to each sample is proportional to the lightness of its frame, with black signifying a weight of 0, white a weight of 1. Those samples belonging to the missing subgroup are assigned the highest weight as they signal from which dataset (training vs. deployment) the bag containing them was drawn from.
  }%
  \label{fig:attn_maps}
\end{figure*}

% \subsection{Qualitative results}\label{sec:qual-results}
\begin{figure*}[tb]
  \centering
  \begin{subfigure}[b]{0.39\textwidth}
    \centering
    \includegraphics[width=\textwidth]{paper3/example_images/fresh-dawn-2179_train_reconstructions_9900.png}
    \caption{
    Different reconstructions on the training set.
    Corresponding to: original, full reconstruction, reconstruction of $z_y$, reconstruction of $z_s$.
    }%
    \label{fig:cmnist-recon-training}
  \end{subfigure}
  % \hfill
  \quad
  \begin{subfigure}[b]{0.39\textwidth}
    \centering
    \includegraphics[width=\textwidth]{paper3/example_images/fresh-dawn-2179_context_reconstructions_9900.png}
    \caption{
    Different reconstructions on the deployment set.
    Corresponding to: original, full reconstruction, reconstruction of $z_y$, reconstruction of $z_s$.
    }%
    \label{fig:cmnist-recon-deployment}
  \end{subfigure}
  \caption{
   Visualization of our method's solutions for the Coloured MNIST dataset, with {\color{purple}purple} as the missing subgroup.
   In each of the subfigures \ref{fig:cmnist-recon-training} and \ref{fig:cmnist-recon-deployment}:
   Column 1 shows the original images from $x$ from the respective set.
   Column 2 shows plain reconstructions generated from $x_\textit{recon}=g(f_y(x), f_s(x))$.
   Column 3 shows reconstruction with zeroed-out $z_s$: $g(f_y(x), 0)$, which effectively visualises $z_y$.
   Column 4 shows the result of an analogous process where $z_y$ was zeroed out instead.
  }%
  \label{fig:cmnist-recon}
\end{figure*}
% Given a learned invariant representation, we can generate a reconstruction to visualise the information contained in it.
% An example of this can be seen in \figref{fig:3dig-examples}.
% This is from our experiment with 3 digits in Coloured MNIST.
% We can clearly see that the reconstructed invariant representation has lost all colour information;
% instead all digits are magenta-coloured, which was the majority colour in the training set.

>>>>

%% Recommended, but optional, packages for figures and better typesetting:
%\usepackage{microtype}
%\usepackage{graphicx}
%\usepackage{wrapfig}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{booktabs} % for professional tables
%\usepackage{amsfonts}       % blackboard math symbols
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{amsthm}
%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%\usepackage{xcolor}         %for colouring the text on colourMNIST
%\usepackage{comment}
%\usepackage{enumitem}
%\newtheorem{prop}{Proposition}
%\newcommand{\Xcal}{\mathcal{X}}
%\newcommand{\Bcal}{\mathcal{B}}
%\newcommand{\Ycal}{\mathcal{Y}}
%\newcommand{\Scal}{\mathcal{S}}
%\newcommand{\Lcal}{\mathcal{L}}
%\newcommand{\Dcal}{\mathcal{D}}
%\newcommand{\ie}{i.\,e.}
%\newcommand{\Ie}{I.\,e.}
%\newcommand{\eg}{e.\,g.}
%\newcommand{\Eg}{E.\,g.}
%% \newcommand{\xmark}{\text{\ding{55}}}
%\usepackage{pifont}% http://ctan.org/pkg/pifont
%\newcommand{\cmark}{\ding{51}}%
%\newcommand{\xmark}{\ding{55}}%

@@

\section{Appendix}\label{sec:zsf-appendix}
\subsection{Results for Adult Income}\label{sec:adult-results}
\begin{figure*}[p]
    \centering
    \includegraphics[width=\columnwidth]{paper3/figures/adult_partial_acc.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_partial_prr.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_partial_tprr.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_partial_tnrr.pdf}
    \caption{%
    Results for the Adult Income dataset with \emph{subgroup bias},
    for the binary classification task of predicting whether an individual earns $>$\$50,000 with a binary subgrouping based on \emph{gender}.
    \texttt{ERM (LD)} refers to a model based on ERM (empirical risk minimisation),
    trained on a \emph{l}abelled \emph{d}eployment set; thus not suffering from bias in the training set.
    \textsc{Top left}: Accuracy.
    \textsc{Top right}: Positive rate ratio.
    \textsc{Bottom left}: True positive rate ratio.
    \textsc{Bottom right}: True negative rate ratio.
    For the \texttt{Ranking} clustering, the clustering accuracy was 69.7\% $\pm$ 0.3\%;
    for \texttt{K-means} it was 43\% $\pm$ 3\%.
    }%
    \label{fig:adult-subgroup-bias}
\end{figure*}
\begin{figure*}[p]
    \centering
    \includegraphics[width=\columnwidth]{paper3/figures/adult_miss_s_acc.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_miss_s_prr.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_miss_s_tprr.pdf}
    \includegraphics[width=\columnwidth]{paper3/figures/adult_miss_s_tnrr.pdf}
    \caption{%
    Results for the Adult Income dataset with a \emph{missing subgroup},
    for the binary classification task of predicting whether an individual earns $>$\$50,000 with a binary subgrouping based on \emph{gender}.
    \texttt{ERM (LD)} refers to a model based on ERM (empirical risk minimisation),
    trained on a \emph{l}abelled \emph{d}eployment set; thus not suffering from bias in the training set.
    \textsc{Top left}: Accuracy.
    \textsc{Top right}: Positive rate ratio.
    \textsc{Bottom left}: True positive rate ratio.
    \textsc{Bottom right}: True negative rate ratio.
    For the \texttt{Ranking} clustering, the clustering accuracy was 60.4\% $\pm$ 0.8\%;
    for \texttt{K-means} it was 44\% $\pm$ 3\%.
    }%
    \label{fig:adult-missing-subgroup}
\end{figure*}
Figures~\ref{fig:adult-subgroup-bias} and \ref{fig:adult-missing-subgroup} show results from our method on the Adult Income dataset \cite{Dua:2017}.
This dataset is a common dataset for evaluating fair machine learning models. 
Each instance in the dataset is described by $14$ characteristics including gender, education, marital status, number of work hours per week among others, along with a label denoting income level ($\geq$\$50K or not). 
We transform the representation into $62$ real and binary features along with the subgroup label $s$. %%%%%%%%CHECK THESE VALUES%%%%%%%%%%%% 
The dataset is naturally imbalanced with respect to gender: 30\% of the males are labelled as earning more than \$50K per year (high income), while only 11\% of females are labelled as such.
For further details on the dataset construction, see section~\ref{ssec:dataset-construction-adult}.
%
Following standard practice in algorithmic fairness, e.g. \citet{zemel2013learning}, we consider gender to be the subgroup label $s$.

@@

As before, \texttt{Ranking}, \texttt{k-means}, \texttt{No bal.}\ and \texttt{Perfect} refer to our method with different procedures for constructing (approximately) perfect bags.
As baseline methods, we have \texttt{ERM} (standard empirical risk minimisation with balanced batches), \texttt{DRO} \citep{HasSriNamLia18}, \texttt{gDRO} \citep{sagawa2019distributionally}
and \texttt{ERM (LD)} which is the same model as \texttt{ERM}, but trained on the labelled deployment set, in addition to the training set.

@@

\paragraph{Adult Income.}\label{ssec:dataset-construction-adult}
For the Adult Income dataset, we do not need to apply any synthetic biasing as the dataset naturally contains some bias wrt $s$. Thus, we instantiate the deployment set as just a random subset of the original dataset. Explicit balancing of the test set is needed to yield meaningful evaluation, however, namely in the penalising of biased classifiers, but need be taken in doing so. Balancing the test set such that
\begin{align}
    |\{x \in X |s=0, y=0\}| &= |\{x \in X |s=1, y=0\}|    \nonumber\\
    \text{and}~|\{x \in X |s=0, y=1\}| &= |\{x \in X |s=1, y=1\}|
\end{align}
where for both target classes, $y=0$ and $y=1$, the proportions of the groups $s=0$ and $s=1$ are made to be the same, is intuitive, yet at the same time precludes sensible comparison of the accuracy/fairness trade-off of the different classifiers.
Indeed, with the above conditions, a majority classifier (predicting all 1s or 0s) achieves comparable accuracy to the fairness-unaware baselines, while also yielding perfect fairness, by construction.
This observation motivated us to devise an alternative scheme, where we balance the test set according to the following constraints
\begin{align}
    & |\{x \in X |s=0, y=0\}| 
    = |\{x \in X |s=0, y=1\}|  \nonumber \\
    = &|\{x \in X |s=1, y=1\}|
    = |\{x \in X |s=1, y=0\}|~.
 \end{align}
That is, all subsets of $\gS \times \gY$ are made to be equally sized. Under this new scheme the accuracy of the the majority classifier is 50\% for the binary-classification task.

\subsection{Optimisation}
\begin{table*}[p]
 \centering
 \caption{Selected hyperparameters for experiments with Coloured MNIST, Adult and CelebA datasets.}
 \label{tab:hparams}
\scalebox{0.65}{
 \begin{tabular}{llll}
 \toprule
 & \textsc{Coloured MNIST} & \textsc{Adult} & \textsc{CelebA}       
 \\ & 2-dig SB / 2-dig MS / 3-dig SB
 \\ \midrule
 Input size  &   $3 \times 32 \times 32$ & $61$ & $3 \times 64 \times 64$ \\  \midrule
 \multicolumn{4}{c}{Autoencoder}                     \\ \midrule
 Levels                      & $4$         & $1$    & $5$\\
 Level depth                 & $2$         & $1$    & $2$\\
 Hidden units / level        & $[32, 64, 128, 256]$ & $[61]$ & $[32, 64, 128, 256, 512]$\\
 Activation                  & GELU        & GELU   & GELU   \\
 Downsampling op.  & Strided Convs. & -- & Strided Convs.\\
 Reconstruction loss         & MSE         & Mixed$^1$  & MSE \\
 Learning rate               & $1 \times 10^{-3}$   & $1 \times 10^{-3}$  & $1 \times 10^{-3}$ \\ \midrule
 \multicolumn{4}{c}{Clustering}                      \\ \midrule
 Batch size                  & $256$      & $1000$  & --\\
 AE pre-training epochs      & $150$        & $100$ & --   \\
 Clustering epochs           & $100$       & $300$  & -- \\
 Self-supervised loss & Cosine + BCE & Cosine + BCE & --\\
 U (for ranking statistics)             & $5$         & $3$     & --    \\   \midrule
 \multicolumn{4}{c}{Distribution Matching}                   \\ \midrule
 Batch size & $1$/$32$/$14$  & $64$   & $32$ \\
 Bag size   & $256$/$8$/$18$ & $32$ & $8$ \\
 Training iterations    & $8\text{k}/8\text{k}/20\text{k}$ & $5\text{k}$ & $15\text{k}$ \\
 Encoding ($z$) size$^2$  & $128$   & $35$  & $128$ \\
 Binarised $z_s$ & {\boxedsymbols ✗}\, / \cmark\, / \cmark & \xmark & \xmark \\
 $y$-predictor weight ($\lambda_1$) & $1$ & $0$ & $1$  \\ 
 $s$-predictor weight ($\lambda_2$) & $1$ & $0$ & $0$  \\ 
 Adversarial weight ($\lambda_3$)   & $1 \times
 10^{-3}$   & $1$   & $1$\\ 
 Stopgradient $\left(\nabla_\theta h_\psi(f_\theta(X^\mathit{dep}))=0\right)$ & \xmark & \cmark & \xmark \\
 \midrule
 \multicolumn{4}{c}{Predictors}   \\ \midrule
 Learning rate  & $3 \times 10^{-4}$ &   $1 \times 10^{-3}$  $ 1 \times 10^{-3}$\\
 \midrule
 \multicolumn{4}{c}{Discriminator}                   \\ \midrule
 Attention mechanism$^3$    & Gated   & Gated & Gated \\
 Hidden units pre-aggregation  & $[256, 256]$  & $[32]$ & $[256, 256]$\\
 Hidden units post-aggregation & $[256, 256]$ & --  & $[256, 256]$ \\
 Embedding dim (for attention) & $32$ & $128$ & $128$ \\
 Activation & GELU & GELU & GELU \\
 Learning rate  & $3 \times 10^{-4}$    & $1 \times 10^{-3}$ & $1 \times 10^{-3}$\\
 Updates / AE update    & $1$  & $3$    & $1$    \\
 \bottomrule
 \multicolumn{4}{l}{\footnotesize $^1$ Cross-entropy is used for categorical features, MSE for continuous features.} \\
  \multicolumn{4}{p{\textwidth}}{\footnotesize $^2$ $|z|$ denotes the combined size of $z_s$ and $z_y$, with the former occupying $\ceil{\text{log}_2(\mathcal{S})}$ dimensions, the latter the remaining } \\
  \multicolumn{4}{p{\textwidth}}{\footnotesize dimensions.} \\
  \multicolumn{4}{p{\textwidth}}{\footnotesize $^3$ The attention mechanism used for computing the sample-weights within a bag. \emph{Gated} refers to gated attention  proposed by
%   \cite{ilse2018attention}, while \emph{SDP} refers to the scaled dot-product attention proposed by \cite{vaswani2017attention}.
  } \\ 
  \multicolumn{4}{p{\textwidth}}{\footnotesize  \cite{ilse2018attention}.} \\ 
 \end{tabular}
}
\end{table*}

@@

Since, by design, we do not have labels for all subgroups the model will be tested on, and bias against these missing subgroups is what we aim to avoid, properly validating, and thus conducting hyperparameter selection for models generally, is not straightforward.
We can use estimates of the mutual information between the learned-representation and $s$ and $y$ (which we wish to minimise w.r.t.\ to the former, maximise w.r.t.\ the latter) to guide the process, though optimising the model w.r.t.\ to these metrics obtained from only the training set does not guarantee generalisation to the missing subgroups.
We can, however, additionally measure the entropy of the predictions on the encoded test set and seek to maximise it across all samples, or alternatively train a discriminator of the same kind used for distribution matching as a measure the shift in the latent space between datasets.
We use the latter approach (considering, the learned distance between subspace distributions, accuracy, and reconstruction loss) to inform an extensive grid-search over the hyperparameter space of our model.

@@

In addition to the losses stated in the distribution matching objective, $\mathcal{L}$, in the main text, we also regularise the encoder by the $\ell^2$ norm of its embedding, multiplied by a small pre-factor, finding this to work better than more complex regularisation methods, such as spectral normalisation \citep{miyato2018spectral}, for stabilising training.

\subsection{Visualisations of results}\label{sec:qual-results}
Figures~\ref{fig:attn_maps} and \ref{fig:cmnist-recon} show some additional visualisations for our results.
For details, see the captions.
\begin{figure*}[tb]
  \centering
    % \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=0.9\textwidth]{paper3/figures/celeba_attn_map.png}
    % \end{subfigure}
    % \hfill
    % \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=0.9\textwidth]{paper3/figures/cmnist_attn_map.png}
    % \end{subfigure}
  \caption{
    Example sample-wise attention maps for bags of CelebA (left) and Coloured MNIST (right) images sampled from a balanced deployment set. The training set is biased according to the SB setting where for CelebA ``smiling females'' constitute the missing source and for Coloured MNIST {\color{purple}purple} fours constitute the missing source. The attention weights are used during the discriminator's aggregation step to compute a weighted sum over the bag. The attention-weight assigned to each sample is proportional to the lightness of its frame, with black signifying a weight of 0, white a weight of 1. Those samples belonging to the missing subgroup are assigned the highest weight as they signal from which dataset (training vs. deployment) the bag containing them was drawn from.
  }%
  \label{fig:attn_maps}
\end{figure*}

% \subsection{Qualitative results}\label{sec:qual-results}
\begin{figure*}[tb]
  \centering
  \begin{subfigure}[b]{0.39\textwidth}
    \centering
    \includegraphics[width=\textwidth]{paper3/example_images/fresh-dawn-2179_train_reconstructions_9900.png}
    \caption{
    Different reconstructions on the training set.
    Corresponding to: original, full reconstruction, reconstruction of $z_y$, reconstruction of $z_s$.
    }%
    \label{fig:cmnist-recon-training}
  \end{subfigure}
  % \hfill
  \quad
  \begin{subfigure}[b]{0.39\textwidth}
    \centering
    \includegraphics[width=\textwidth]{paper3/example_images/fresh-dawn-2179_context_reconstructions_9900.png}
    \caption{
    Different reconstructions on the deployment set.
    Corresponding to: original, full reconstruction, reconstruction of $z_y$, reconstruction of $z_s$.
    }%
    \label{fig:cmnist-recon-deployment}
  \end{subfigure}
  \caption{
   Visualisation of our method's solutions for the Coloured MNIST dataset, with {\color{purple}purple} as the missing subgroup.
   In each of the subfigures \ref{fig:cmnist-recon-training} and \ref{fig:cmnist-recon-deployment}:
   Column 1 shows the original images from $x$ from the respective set.
   Column 2 shows plain reconstructions generated from $x_\textit{recon}=g(f_y(x), f_s(x))$.
   Column 3 shows reconstruction with zeroed-out $z_s$: $g(f_y(x), 0)$, which effectively visualises $z_y$.
   Column 4 shows the result of an analogous process where $z_y$ was zeroed out instead.
  }%
  \label{fig:cmnist-recon}
\end{figure*}
% Given a learned invariant representation, we can generate a reconstruction to visualise the information contained in it.
% An example of this can be seen in \figref{fig:3dig-examples}.
% This is from our experiment with 3 digits in Coloured MNIST.
% We can clearly see that the reconstructed invariant representation has lost all colour information;
% instead all digits are magenta-coloured, which was the majority colour in the training set.
